---
output:
  word_document: default
  html_document: default
---
```{r}
library(tidyverse)
library(cluster)
library(factoextra)
library(dendextend)
```

```{r}
library(readr)
trucks <- read_csv("trucks.csv")
```
```{r}
trucks = trucks %>% drop_na()
```

Task 1. 
```{r}
library(ggplot2)
ggplot(trucks, aes(x = Distance, y = Speeding)) +
  geom_point()
```
Based on the scatterplot, there does seem to be some natural clustering around speeding when traveling 25-75 miles and another natural cluster when traveling between 150-200 miles. Both have large clusters that fall below 25% of time spent traveling but the shorter distance seems to have another cluster above 25% of the time as well. 

Task 2. 
```{r}
trucks2 = trucks %>% select(-Driver_ID)
trucks2 = as.data.frame(scale(trucks2))
summary(trucks2)
```

Task 3.
```{r}
set.seed(64)
clusters1 <- kmeans(trucks2, 2)
fviz_cluster(clusters1, trucks2)
```
Both clusters seem to be independent of each other and similar to the scatterplot I graphed earlier in task 2. Cluster 2 is more scattered than cluster 1.

Task 4. 
```{r}
set.seed(64)
fviz_nbclust(trucks2, kmeans, method = "wss")
```
```{r}
set.seed(64)
fviz_nbclust(trucks2, kmeans, method = "silhouette")
```
Yes, there is a consensus between these 2 methods to use k=4 clusters. 

Task 5. 
```{r}
set.seed(64)
clusters2 <- kmeans(trucks2, 4)
fviz_cluster(clusters2, trucks2)
```

Task 6.
Clusters 1 and 3 seem to have similar characteristics of speeding, distance-wise. Both groups, who seem to not travel over 100 miles, tend to speed under 50% of the time. Clusters 2 and 4 also share similar characteristics of speeding distance-wise. While the two clusters do not overlap, a larger group in each cluster shows that the longer distances they have traveled (over 100 miles), the longer time they tend to drive over the speed limit. Cluster 4 looks to be the biggest one but, similar to clusters 1 and 3, they tend to not speed more than 50% of the time. Cluster 2 on the other hand is a bit more spread out, so less drivers seem to be grouped into this cluster, but these drivers tend to speed 50-100% of the time they are driving. 

Task 7.
```{r}
bball = read_csv("kenpom20.csv")
```
```{r}
bball2 = bball %>% select(-TeamName)
bball2 = as.data.frame(scale(bball2))
```

```{r}
set.seed(123)
fviz_nbclust(bball2, kmeans, method = "wss")
```
```{r}
set.seed(123)
fviz_nbclust(bball2, kmeans, method = "silhouette")
```
There's not a clear consensus between the 2 methods to use k = 2, and the wss method doesn't seem to depict that very clearly like the silhouette method does. I think the agreement between the two methods is that a big shift happened between k 1 and 2 based on a large sum, so k = 2 could be used. But k = 4 could be another solution based on the "elbows" of both graphs. 

Task 8.
```{r}
set.seed(1234)
clusters3 <- kmeans(bball2, 4)
fviz_cluster(clusters3, bball2)
```

Task 9.
```{r}
bball2 = bball2 %>% mutate(clusternum = clusters3$cluster)
```

```{r}
ggplot(bball2, aes(x = AdjOE, y = AdjDE, color = factor(clusternum))) +
  geom_point()
```
The majority of the clusters, especially clsuters 1 and 4, score about 0 - 1 points per 100 possessions on both the offensive and defensive ends. Cluster 2 tends to score more points when playing offense while cluster 3 is even more spread out and tends to score more points playing defense. 
